---
title: "Data Analysis of NYCFlights"
author: "Preeti Jha"
date: "2/21/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
```{r}
library(tidyverse)
library(skimr)
```

```{r warning=FALSE}
library(nycflights13)
data(flights)
head(flights)
```
 
 
 
 
# Number of NA in dep_time is the number of cancelled flights
```{r warning=FALSE}
flights%>%
  filter(is.na(dep_time))%>%
  count()
```
# Maximum number of NA or missing values of flights is in the month of february so highest number of flight cancellations happened in *February* and lowest number of cancellations happened in the month of *november*
```{r warning=FALSE}
flights %>%select(month,dep_time)%>%
  group_by(month)%>%
  skim()%>%
  arrange(desc(n_missing))

```
          
# Tail number of the plane with the most flights from NYC is *N725MQ*.
```{r warning=FALSE}
#names(flights)
flights %>% select(tailnum)%>%
  group_by(tailnum)%>%
  count(count = n())%>%
  arrange(desc(count))
```
# Visualizing  number of trips per week over the year
```{r warning=FALSE}

flights %>% select(year, month, day, tailnum)%>%
  filter(tailnum == "N725MQ")%>%
  group_by(year,month) %>%
  count()

```


```{r warning=FALSE}
flights %>% filter(tailnum == "N725MQ") %>%
  group_by(month) %>%
  count() %>%
  ggplot(aes(x = month, y = n)) +
  geom_point() +
  geom_line()+
  ggtitle("Number of trips per week")+xlab("week")+
  ylab("Flights per week")


```

# Planes Dataset

```{r warning=FALSE}
head(planes)
names(planes)
#?planes after seeing through help file of planes data it seems that year coloumn is the year of manufacturing of planes
```

# Flights Dataset
```{r warning=FALSE}
#Renaming departure year as year
flights_2 <- flights%>%
 rename(dep_year = "year")
head(flights_2)

```
# *N381AA* is the oldest plane that flew from NYC in year 2013

```{r warning=FALSE}
#Using join to join flights_2 and planes table to know dep_year ,manufacture year and tailnum
flights_2%>%
  left_join(planes,by= c("tailnum" = "tailnum"))%>%
  select(dep_year,year,tailnum)%>%
  arrange(-desc(year))%>%
  distinct()%>%
  head(1)
```
There are 3322 airplanes that flew from New york City that are included in planes table.

```{r warning=FALSE}
flights_2%>%
      right_join(planes, by = c("tailnum"="tailnum"))%>%
  group_by(tailnum) %>%
   slice(1) %>%
      ungroup() %>%
      count()
```

# There are *70* date missing in planes
```{r warning=FALSE}
 planes%>%
      summarise(missing_date=sum(is.na(year)))
```



# The five most common manufacturers in planes dataset:
```{r warning=FALSE}
planes %>%
  group_by(manufacturer) %>%
  count(most_common_manuf = n())%>%
  arrange(desc(most_common_manuf))%>%
  head(5)

 
```

# The distribution of manufacturer changed over time 

```{r warning=FALSE}
planes_2 <- planes%>%mutate(manufacturer_2 = ifelse(manufacturer %in% c("BOEING", "AIRBUS INDUSTRIE", "BOMBARDIER INC", "AIRBUS", "EMBRAER") ,manufacturer,"Others"))

flights %>%
  inner_join(planes_2, by = c("tailnum" = "tailnum")) %>%
  select(manufacturer_2, time_hour) %>%
  group_by( manufacturer_2,time_hour) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = time_hour, y = n, color = manufacturer_2)) +
  geom_line() + 
  ggtitle("Distribution of Top 5 Most Common Manufacturers", subtitle = "Airplanes flying from NYC in 2013")
```



